---
output: md_document
---

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/umich-cphds/snif.svg?branch=master)](https://travis-ci.org/umich-cphds/snif)
<!-- badges: end -->

# Introduction
`snif` is a R package that implements "Selection of Nonlinear Interactions by a
Forward Stepwise Algorithm". `snif` (the function) is a a forward stepwise
algorithm that takes nonlinearity and interactions into account.

`snif` is currently in the middle of being tested and polished, and as such it
is *BETA* software.

# Installation
`snif` requires the packages `purrr` and `rlang` to be installed, and it is a
good idea to update them to make sure the package works correctly.


```{r, eval = F}
install.packages(c("rlang", "purrr"))

# install.packages("devtools")
devtools::install_github("umich-cphds/snif")
```
# Example
This section is based off the snif documentation.

The first argument to `snif` is a R formula that defines the initial model `snif`
will use during the forward stage of the algorithm. the second argument, `df`,
is a R data.frame, which is used for scoring, and the columns of `df` are the
possible variables that `snif` will consider. In other words, if there variables
in `df` that you don't want snif to consider, you should subset `df` first.

`snif` supports binary variables by setting `type = "logistic"`. Then `snif`
will use a `glm` to fit the data instead of a `lm`. The default for `type` is
linear.

`snif` contains a toy dataset, `snif.df` which is used in this example. It has
500 rows and 11 columns.

```{r}
library(snif)

names(snif.df)
```

In this data.frame, the response, `y`, is continuous so we will use a linear
model by setting `type = "linear"`. We would also like to use BIC scoring to
determine which variable to model to the model, so we set `score = "BIC"`. For
the initial formula to fit, we do not know much about this data so we will start
with the null model, `y ~ NULL`.

```{r}
snif.out <- snif(formula = y ~ NULL, df = snif.df, type = "linear", score = "BIC")

summary(snif.out)
```

Alternatively, we can use a non null initial model
```{r}
snif.out <- snif(formula = y ~ V2, df = snif.df, type = "linear",
                   score = "BIC")

summary(snif.out)
```

`snif` also supports interaction terms via `:` and basis expansions via `bs`.

```{r}
snif.out <- snif(formula = y ~ bs(V2) + V2:V4, df = snif.df, type = "linear",
                   score = "BIC")
```
Note that `bs` implies that both `V2` and the nonlinear expansion of `V2`,
`bs(V2)[,-1]` are both included in the model. That is the following two
formulas are equivalent: `y ~ bs(V2)` and `y ~ V2 + bs(V2)[,-1]`
If you wish for only the non linear part of `V2` to be included in `snif`, you
need to drop the linear term from the expansion, that is, `y ~ bs(V2)[, -1]`.
Finally, you may specify the degree of the expansion via the degree parameter in
`snif`.

More detailed documentation is available on the `snif` help page, which
provides rules for building initial `snif` formulas, and documents other optional
parameters not mentioned here.

# Reference
Narisetty, Naveen N. and Mukherjee, Bhramar and Chen, Yin-Hsiu and Gonzalez,
Richard and Meeker, John D. Selection of nonlinear interactions by a
forward stepwise algorithm: Application to identifying environmental
chemical mixtures affecting health outcomes. Statistics in Medicine.
2019;38(9):1582-1600. \href{https://doi.org/10.1002/sim.8059}{10.1002/sim.8059}

# Contact
Please email Alex Rix (alexrix@umich.edu) if you have any questions,
comments, or possible bugs to report. You may also open up an issue on github.
